\IfFileExists{elsarticle.cls}{%
\documentclass[preprint,12pt]{elsarticle}
\newif\ifuseelsevier
\useelseviertrue
}{%
\documentclass[12pt]{article}
\newif\ifuseelsevier
\useelsevierfalse
}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{hyperref}
\hypersetup{hidelinks}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\graphicspath{{figures/}}
\emergencystretch=3em

\begin{document}
\ifuseelsevier
\begin{frontmatter}
\title{When Does Reinforcement Learning Improve Supply Chain Resilience? Action-Space Alignment in a Garrido-Grounded DES Benchmark}
\begin{abstract}
Supply chain resilience studies increasingly combine simulation and reinforcement learning, yet it remains unclear when neural control improves resilience beyond strong static baselines. We build a Garrido-grounded discrete-event simulation benchmark of a 13-operation military food supply chain and evaluate PPO under common random numbers, dense static dispatch frontiers, and workbook-faithful Garrido/Excel resilience metrics. Under Track A, which preserves the original buffer/shift decision family, learned policies fail to convert the small measured headroom over the best static policy. Under Track B, a DES-preserving extension that adds downstream dispatch control at Op10 and Op12, PPO improves Excel ReT over the best dense downstream-dispatch static comparator by +0.000438 across 10 training seeds, with fully positive seed-clustered intervals and matching gains in flow fill, service-loss AUC, backlog, and recovery tails. The gain is not explained by privileged simulator fields: masked PPO retains most of the advantage, while a regime-conditioned non-learning lookup table remains far below PPO. Mechanism checks and local static bounds indicate that the result is tied to exposing the downstream recovery bottleneck, not to action-space size alone or an untested full 8D static frontier. We conclude that RL value is observed in this benchmark when the controllable interface reaches the binding operational bottleneck and the resulting headroom is dynamically convertible into adaptive recovery performance.
\end{abstract}
\begin{keyword}
Supply chain resilience \sep reinforcement learning \sep discrete-event simulation \sep military supply chain \sep bottleneck alignment
\end{keyword}
\end{frontmatter}
\else
\title{When Does Reinforcement Learning Improve Supply Chain Resilience? Action-Space Alignment in a Garrido-Grounded DES Benchmark}
\author{}
\date{}
\maketitle
\begin{abstract}
Supply chain resilience studies increasingly combine simulation and reinforcement learning, yet it remains unclear when neural control improves resilience beyond strong static baselines. We build a Garrido-grounded discrete-event simulation benchmark of a 13-operation military food supply chain and evaluate PPO under common random numbers, dense static dispatch frontiers, and workbook-faithful Garrido/Excel resilience metrics. Under Track A, which preserves the original buffer/shift decision family, learned policies fail to convert the small measured headroom over the best static policy. Under Track B, a DES-preserving extension that adds downstream dispatch control at Op10 and Op12, PPO improves Excel ReT over the best dense downstream-dispatch static comparator by +0.000438 across 10 training seeds, with fully positive seed-clustered intervals and matching gains in flow fill, service-loss AUC, backlog, and recovery tails. The gain is not explained by privileged simulator fields: masked PPO retains most of the advantage, while a regime-conditioned non-learning lookup table remains far below PPO. Mechanism checks and local static bounds indicate that the result is tied to exposing the downstream recovery bottleneck, not to action-space size alone or an untested full 8D static frontier. We conclude that RL value is observed in this benchmark when the controllable interface reaches the binding operational bottleneck and the resulting headroom is dynamically convertible into adaptive recovery performance.
\end{abstract}
\fi

\input{sections/01_introduction.tex}
\input{sections/02_related_work.tex}
\input{sections/03_methodology.tex}
\input{sections/04_results.tex}
\input{sections/05_discussion.tex}
\input{sections/06_conclusion.tex}

\section*{CRediT authorship contribution statement}
[TO COMPLETE before submission: author names and roles, e.g.
Conceptualization; Methodology; Software; Validation; Formal analysis;
Writing -- original draft; Writing -- review and editing; Supervision.]

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial
interests or personal relationships that could have appeared to
influence the work reported in this paper.

\section*{Funding}
[TO COMPLETE: funding statement, or ``This research did not receive any
specific grant from funding agencies in the public, commercial, or
not-for-profit sectors.'']

\section*{Data and code availability}
All simulation code (DES backbone, Gymnasium wrappers, training and
evaluation scripts), the dense static frontier definitions, the frozen
evidence bundles behind every table and figure, and the figure-generation
pipeline are openly available in the project repository
(\texttt{github.com/Thom-320/scres-ia}, branch
\texttt{codex/garrido-replication-experiments}). The original
Garrido-Rios workbook artifacts used for the row-level metric audit are
available from the cited thesis repository (WRAP, University of Warwick)
and from the corresponding author on reasonable request.

\section*{Acknowledgments}
[TO COMPLETE: acknowledge A. Garrido for the thesis materials and
workbook artifacts if appropriate, plus any colleagues/reviewers.]

\appendix

\section{Comparator-selection robustness}\label{appendix-robustness}

Because the headline comparator (the strongest of 147 statics) is
selected on the same evaluation episodes used to score PPO, comparator
selection could in principle inflate the reported delta (a
winner's-curse concern). Table~\ref{tab:top12_robustness} recomputes the
canonical paired delta against each of the twelve strongest static
comparators individually. All twelve deltas are positive with confidence
intervals excluding zero, and the delta \emph{grows} down the ranking ---
the opposite of the pattern selection bias would produce. Comparator
selection cannot flip the sign of the result.

\begin{table}[htbp]
\centering
\caption{PPO Excel ReT delta against each of the top-12 static dispatch
comparators (60 CRN pairs each; bootstrap CI95). Every pairwise delta is
positive with a fully positive interval.}
\label{tab:top12_robustness}
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{clccc}
\toprule
Rank & Static policy & Static Excel ReT & $\Delta$ & CI95 \\
\midrule
1 & S2\_op10\_2.00\_op12\_1.50 & 0.005466 & +0.000426 & [+0.000388, +0.000463] \\
2 & S2\_op10\_1.25\_op12\_2.50 & 0.005466 & +0.000427 & [+0.000380, +0.000475] \\
3 & S3\_op10\_2.50\_op12\_2.50 & 0.005465 & +0.000428 & [+0.000389, +0.000467] \\
4 & S3\_op10\_1.25\_op12\_2.00 & 0.005450 & +0.000443 & [+0.000396, +0.000494] \\
5 & S2\_op10\_1.00\_op12\_2.00 & 0.005445 & +0.000447 & [+0.000404, +0.000489] \\
6 & S3\_op10\_1.25\_op12\_1.50 & 0.005442 & +0.000450 & [+0.000404, +0.000497] \\
7 & S3\_op10\_2.00\_op12\_2.50 & 0.005442 & +0.000451 & [+0.000411, +0.000491] \\
8 & S3\_op10\_1.00\_op12\_2.50 & 0.005441 & +0.000452 & [+0.000410, +0.000493] \\
9 & S2\_op10\_2.50\_op12\_1.25 & 0.005438 & +0.000454 & [+0.000422, +0.000488] \\
10 & S3\_op10\_2.50\_op12\_2.00 & 0.005438 & +0.000455 & [+0.000424, +0.000487] \\
11 & S2\_op10\_2.50\_op12\_2.00 & 0.005433 & +0.000460 & [+0.000419, +0.000504] \\
12 & S2\_op10\_1.50\_op12\_1.50 & 0.005428 & +0.000465 & [+0.000422, +0.000507] \\
\bottomrule
\end{tabular}
\end{table}

\bibliographystyle{elsarticle-num}
\bibliography{references}

\end{document}
