\hypertarget{2-related-work}{%
\section{Related Work}\label{2-related-work}}

\hypertarget{21-scres-measurement-and-adaptive-capability}{%
\subsection{SCRES measurement and adaptive capability}\label{21-scres-measurement-and-adaptive-capability}}

Supply chain resilience (SCRES) is commonly defined around the ability of
a supply chain to prepare for, respond to, and recover from disruptions
\cite{ponomarov2009understanding}. More recent theory distinguishes
resilience as a system property from resilience as a dynamic capability
that must be renewed through experience and organizational learning
\cite{wieland2021two,levitt1988organizational}. Quantitative SCRES
research has consequently moved beyond binary disruption survival toward
metrics for recovery speed, service continuity, tail performance, and
system viability \cite{hosseini2019review}. This paper follows that
measurement tradition: reinforcement learning is evaluated on
Garrido/Excel ReT and order-level recovery metrics, not on the training
reward alone.

\hypertarget{22-garrido-grounded-des-for-resilience}{%
\subsection{Garrido-grounded DES for resilience}\label{22-garrido-grounded-des-for-resilience}}

Garrido-Rios~\cite{garrido2017} provides an unusually useful starting
point for AI-enabled SCRES because the military food supply chain is
already modeled as a detailed 13-operation simulation with resilience
indices, risk families, and static policy comparisons. Garrido et
al.~\cite{garrido2024ijpr} extend that lineage through a
Cobb--Douglas-style factory resilience index, while Garrido et
al.~\cite{garrido2024iccl} explicitly identify learning algorithms as a
future path for simulation-based SCRES models. The present study answers
a narrower and more testable version of that agenda: not whether AI can
be attached to a DES, but whether learned closed-loop control can beat
strong static resilience baselines when the controllable decision surface
does or does not reach the operational recovery bottleneck.

\hypertarget{23-rl-for-supply-chain-operational-control}{%
\subsection{RL for supply-chain operational control}\label{23-rl-for-supply-chain-operational-control}}

Reinforcement learning has become an established paradigm for adaptive
decision-making in supply chain management. Yan et al.~\cite{yan2022}
and Rolf et al.~\cite{rolf2023} review applications across inventory
ordering, vehicle routing, warehouse operations, and multi-echelon
coordination. Boute et al.~\cite{boute2022} provide a design roadmap for
deep RL in inventory control, and Powell~\cite{powell2022} cautions that
neural value-function approximation is only one policy class among
several structured alternatives. Empirical inventory-control studies
reinforce that warning: deep RL can perform well, but strong base-stock,
teacher, and heuristic policies remain difficult comparators
\cite{gijsbrechts2022,demoor2022,dehaybe2024,geevers2024,temizoz2025,vanvuchelen2025}.
Our contribution is therefore not the generic application of PPO to a
supply chain. It is a controlled benchmark showing that action-surface
design can decide whether any learned policy has a useful recovery
frontier to exploit.

\hypertarget{24-des-ml-and-ai-scres}{%
\subsection{DES+ML and AI-SCRES workflows}\label{24-des-ml-and-ai-scres}}

Hybrid simulation and machine-learning workflows are now common in
supply-chain analysis. Kogler and Maxera~\cite{kogler2026} review 72
papers and 43 related studies combining discrete simulation with machine
learning, while Badakhshan et al.~\cite{badakhshan2024} show that many
simulation-ML workflows emphasize prediction or classification rather
than optimization. In resilience-oriented AI, Ding et al.~\cite{ding2026}
provide the closest resilience-oriented multi-agent RL comparator: MAPPO
agents for suppliers, manufacturers, and distributors reconfigure a
supply-chain system-of-systems through filling, repairing, and
recruiting actions, evaluated on a composite of complex-network
statistics and benchmarked against QMIX and MADDPG. Our study addresses
a different benchmark question at a different level: whether learned
\emph{operational} control improves a Garrido-grounded DES beyond dense
static and heuristic baselines under common-random-number evaluation,
scored on an order-level recovery-process metric rather than network
statistics, and with the training reward deliberately separated from the
reported resilience index. Elsewhere in this stream,
Bussieweke et al.~\cite{bussieweke2024recovery} combine system dynamics
and RL for recovery policies, Kim~\cite{kim2024} studies cooperative
MARL for inventory transshipment, Liu et al.~\cite{liu2025pom} apply
HAPPO to decentralized multi-echelon inventory, and Kotecha and del Rio
Chanona~\cite{kotecha2025} combine graph neural networks with MAPPO.
Other AI-SCRES studies focus on prediction, classification, or
explanation rather than online control
\cite{rajagopal2024,rezki2024,ordibazar2022}. These streams establish
the relevance of AI for SCRES, but they rarely test learned control
against dense static frontiers on a thesis-grounded resilience metric.

\begin{table}[htbp]
\centering
\caption{Positioning of the present benchmark relative to adjacent literatures. The table is a conceptual positioning matrix over representative papers discussed in this section, not a systematic coding of the full literature.}
\label{tab:literature_gap}
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{p{0.24\linewidth}>{\centering\arraybackslash}p{0.125\linewidth}>{\centering\arraybackslash}p{0.125\linewidth}>{\centering\arraybackslash}p{0.125\linewidth}>{\centering\arraybackslash}p{0.125\linewidth}>{\centering\arraybackslash}p{0.125\linewidth}}
\toprule
Stream & Validated DES & SCRES metric & Dense static frontier & Bottleneck action test & Priv.-obs.\ defense \\
\midrule
DES SCRES models & Yes & Yes & Sometimes & Rare & No \\
RL inventory/control & Rare & No & Sometimes & Rare & Rare \\
DES+ML supply-chain workflows & Sometimes & Rare & Rare & Rare & Rare \\
AI-SCRES / MARL reconfiguration & Sometimes & Sometimes & Rare & Strategic & Rare \\
This paper & Yes & Yes & Yes & Yes & Yes \\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{25-gap-action-space-alignment}{%
\subsection{Gap: action-space alignment and bottleneck authority}\label{25-gap-action-space-alignment}}

This literature also motivates our evaluation standard. Deep RL for
inventory control repeatedly finds that strong structured policies and
classical heuristics remain hard to beat: Gijsbrechts et
al.~\cite{gijsbrechts2022} report that DRL matches but often does not
decisively exceed state-of-the-art heuristics, Stranieri et
al.~\cite{stranieri2024} find PPO excels in complex, variable
pharmaceutical settings without dominating every classical policy, and
structure-informed designs~\cite{maggiar2025} argue that exploiting known
problem structure matters more than generic architecture. The relevant
question for a learned policy is therefore not whether it beats a naive
static rule, but whether it beats a densely sampled family of strong
static and heuristic policies under paired stochastic evaluation --- the
standard adopted throughout this paper.

The gap is therefore conditional rather than categorical. Prior work
shows that RL can be used in supply chains and that DES can represent
SCRES dynamics. What remains under-tested is whether the learned policy's
action space actually reaches the constraint that governs recovery.
Theory-of-Constraints reasoning treats the binding bottleneck as the
lever that determines system throughput \cite{goldratt1984goal}, and the
factory-physics tradition formalizes why buffer and capacity
interventions away from the binding constraint yield sharply diminishing
returns \cite{hopp2011factory}; RL
agent design determines whether that lever is even available to the
policy. This paper connects the two: Track A tests learned control when
the original buffer/shift surface leaves the downstream recovery
bottleneck mostly fixed, while Track B tests a DES-preserving extension
that exposes downstream dispatch authority. The contribution is a
benchmark-centered diagnosis of when AI-enabled SCRES control has a
convertible recovery frontier and when it does not.
