\section{Decision-right eligibility framework}\label{sec:framework}

The methodological premise is that an available action is not automatically a useful
sequential decision. Let $S_t$ denote the physical state, $O_t$ the information available to
an operator, $A_t$ an admissible action, and $J$ an episode-level resilience outcome. For a
fixed horizon $H$, define $Q_H(s,a)=\mathbb{E}[J_{t:t+H}\mid S_t=s,A_t=a]$. The value of
perfect information over the best constant action is
\begin{equation}
H_{\mathrm{PI}}=\mathbb{E}_{S}\left[\max_a Q_H(S,a)\right]
-\max_a\mathbb{E}_{S}[Q_H(S,a)].
\end{equation}
The corresponding observable value replaces the true state by the operator's information:
\begin{equation}
H_{\mathrm{obs}}=\mathbb{E}_{O}\left[\max_a\mathbb{E}[Q_H(S,a)\mid O]\right]
-\max_a\mathbb{E}_{S}[Q_H(S,a)].
\end{equation}
These quantities separate two questions often conflated in simulation--RL studies. An oracle
may profit from knowing a future disruption, while a deployable controller cannot infer the
relevant action ranking from contemporaneous observations. In that case $H_{\mathrm{PI}}>0$
but $H_{\mathrm{obs}}\approx0$; further optimization of a learner cannot manufacture the
missing information.

We operationalize eligibility as a six-stage ladder. \emph{Physical liveness} requires that
two admissible actions produce different physical trajectories under exact replay.
\emph{Ranking diversity} requires that the preferred action changes across a material fraction
of states; otherwise a constant posture is sufficient. \emph{Oracle headroom} requires a
positive and practically material paired outcome over a same-contract static comparator.
\emph{Observable convertibility} requires a policy using only deployable fields to recover
value in rollout, rather than merely classify oracle actions accurately. \emph{Resource
equivalence} requires the apparent gain to survive equal decision rights and an explicit
budget envelope. Finally, \emph{robustness} requires confirmation on frozen, previously
unopened tapes and guardrails for service, lost orders, and tail outcomes. Failure at an
earlier stage blocks claims at every later stage.

The comparator contract is central. If an RL policy controls eight variables while the static
frontier varies only three, the contrast estimates a mixture of adaptation and additional
authority. The correct static bar is the strongest implementable constant or open-loop doctrine
with the same action rights and no greater resource use. Calibration selects the comparator
once; evaluation does not reselect it per tape. Common random numbers then pair demand,
processing, repair, and disruption realizations across policies. This design turns differences
into action effects rather than differences in exogenous luck.

Observable conversion is evaluated by rollout value, not action accuracy. Suppose an oracle
labels action $a^*(s)$. A classifier minimizes $\mathbf{1}\{\hat a\ne a^*\}$, assigning equal
weight to harmless and catastrophic mistakes. The decision-relevant loss is instead regret,
\begin{equation}
\mathcal{L}_{\mathrm{regret}}(s,\hat a)=Q_H(s,a^*)-Q_H(s,\hat a),
\end{equation}
and the decisive endpoint remains episode performance against the frozen comparator. This
distinction became empirically important: a high oracle-action classification rate did not
translate into a useful sequential policy.

The framework also separates metric constructs. Order-level ReT treats continuity and lateness
per order; quantity-weighted ReT weights the same order scores by ration demand; service-loss
measures physical unmet mass; worst-node fill exposes spatial sacrifice; and a multiplicative
Cobb--Douglas index encodes a preference over inventory, backlog, time, capacity, and resource
dimensions. None is a neutral synonym for ``resilience.'' A policy-ranking reversal across
these measures is reported as construct sensitivity, not resolved by choosing retrospectively
the metric on which a learner wins.

Preregistration makes the ladder falsifiable. Each program freezes its action contract, tapes,
estimand, comparator, practical threshold, and stop rule before observing the relevant data.
A failed gate is retained as a boundary result. New physics or a new information contract
requires a separately named study and cannot alter the earlier verdict. This discipline is
particularly important when environment design itself can create or destroy learnability.

