\hypertarget{3-methodology}{%
\section{3. Methodology}\label{3-methodology}}

Status: draft prose base aligned to the frozen paper contracts.

\hypertarget{31-mfsc-discrete-event-simulation}{%
\subsection{3.1 MFSC discrete-event
simulation}\label{31-mfsc-discrete-event-simulation}}

The experimental platform is a 13-operation military food supply chain
reconstructed in Python/SimPy from Garrido-Rios' thesis. The model
preserves the original operational structure rather than replacing it
with a stylized benchmark environment. This matters for the paper's
positioning: the contribution is not a generic RL sandbox, but a
thesis-grounded operational model whose buffers, transport stages, and
disruption processes already carry domain meaning.

The DES retains hourly granularity where the original supply chain
requires sub-day behavior, especially around the assembly-line
operations. This avoids collapsing operational effects that would
disappear under daily aggregation, including short breakdowns and
shift-sensitive capacity changes. The warm-up regime also follows the
validated thesis lineage, with the first full batch reaching the
downstream chain after roughly \texttt{838.8} simulated hours. The RL
layer therefore sits on top of a model that is already operationally
constrained before learning begins.

The simulation backbone is not treated as a moving target. Deterministic
and stochastic checks remain anchored to the thesis reference tables,
and the benchmark story is built on the audited post-alignment DES
rather than on earlier historical artifacts. This allows the paper to
claim benchmark fidelity without claiming that the model is a universal
representation of all supply chains.

\hypertarget{32-control-formulation}{%
\subsection{3.2 Control formulation}\label{32-control-formulation}}

The learning problem is defined as weekly operational control over the
DES. At each decision epoch, the agent observes a state vector derived
from the simulation, selects a continuous action, and receives a
resilience-oriented reward. Decisions are taken every \texttt{168}
simulated hours, which matches the weekly planning cadence already
embedded in procurement, disruption, and operating rhythms. For
manuscript language, the problem should be described as an augmented
observed-state control problem, or equivalently a practical
\texttt{POMDP}, rather than as a fully sufficient MDP.

The paper uses two control contracts. \texttt{Track\ A} is the
thesis-faithful benchmark family. It uses observation version
\texttt{v1} with \texttt{15} dimensions and a \texttt{5D} action
contract covering upstream quantity and reorder multipliers together
with assembly-shift selection. This is the negative paper lane: it asks
whether RL can beat strong static policies when control remains limited
to the original scope.

\texttt{Track\ B} is a separate research lane rather than a retrofit of
\texttt{Track\ A}. It uses observation version \texttt{v7} with
\texttt{46} dimensions and a \texttt{7D} action contract. The first five
actions keep the Track A controls, while the sixth and seventh actions
add downstream dispatch multipliers at \texttt{Op10} and \texttt{Op12}.
The paper's central question is whether this minimal extension changes
the benchmark outcome by exposing the operational bottleneck that Track
A leaves mostly outside the controllable interface.

\hypertarget{33-reward-and-reporting-metrics}{%
\subsection{3.3 Reward and reporting
metrics}\label{33-reward-and-reporting-metrics}}

The main paper lane uses \texttt{ReT\_seq\_v1} with
\texttt{ret\_seq\_kappa\ =\ 0.20} as the training reward. The reason for
centering this reward is not that it is the only reward ever tested in
the repository, but that it is the best-aligned resilience-index family
that remains smooth enough for policy-gradient learning while still
preserving a visible link to the thesis resilience framing. In the
current paper story, reward design and control-contract design are
discussed jointly: Track A shows that a reasonable resilience-aware
reward is not enough when the controllable action scope misses the
binding downstream constraint, whereas Track B shows that the same
reward can support strong learning once the action contract is repaired.

The paper-facing comparison should therefore prioritize service and
resilience outcomes rather than raw reward totals across incompatible
families. The primary metrics for the main claim are
\texttt{fill\_rate}, \texttt{backorder\_rate},
\texttt{order\_level\_ret\_mean}, and
\texttt{terminal\_rolling\_fill\_rate\_4w}. Secondary audit metrics such
as \texttt{ret\_thesis\_corrected\_total}, \texttt{ret\_unified\_total},
and the order-case shares remain useful for interpretation, but they
should not replace the main service/resilience metrics when the paper
compares policies across benchmark families.

One language rule should stay explicit in the final manuscript:
\texttt{reward\_total} is only comparable within a shared reward family.
Whenever the paper compares Track A and Track B, or different reward
families inside Track A, the argument should rely on service and
resilience metrics rather than on raw cumulative reward.

\hypertarget{34-experimental-protocol}{%
\subsection{3.4 Experimental protocol}\label{34-experimental-protocol}}

The experimental design contains two benchmark families and one matched
ablation. \texttt{Track\ A} is the thesis-faithful benchmark lane. It
retains the validated DES backbone, the original operational scope, and
strong static baselines as the main comparators. Learned policies are
reported honestly as negative evidence when they remain below the
strongest static baseline; this includes PPO in the frozen
\texttt{ReT\_seq\_v1} lane and RecurrentPPO in the corresponding
recurrent comparator lane.

\texttt{Track\ B} is the positive paper lane. Its frozen contract is
\texttt{env\_variant="track\_b\_adaptive\_control"},
\texttt{reward\_mode="ReT\_seq\_v1"}, \texttt{ret\_seq\_kappa=0.20},
\texttt{observation\_version="v7"},
\texttt{action\_contract="track\_b\_v1"},
\texttt{risk\_level="adaptive\_benchmark\_v2"},
\texttt{year\_basis="thesis"}, \texttt{step\_size\_hours=168}, and
\texttt{stochastic\_pt=True}. The production benchmark is trained for
\texttt{500000} timesteps with seeds \texttt{11\ 22\ 33\ 44\ 55} and
evaluated over \texttt{20} episodes per seed. Static comparators are
\texttt{s2\_d1.00}, \texttt{s3\_d1.00}, and \texttt{s3\_d2.00}.

To separate the action-contract argument from broader environment
changes, the paper also uses a matched \texttt{5D\ vs\ 7D} ablation.
Within that ablation family, the observation version, reward family,
risk profile, and training budget are held fixed while only the action
contract changes. This is the cleanest evidence available inside the
current repo for the claim that downstream control coverage is
materially responsible for the Track B gain. The manuscript should still
state the limitation clearly: the ablation is supportive causal evidence
inside the frozen benchmark family, not universal proof outside it.
