\hypertarget{4-results}{%
\section{4. Results}\label{4-results}}

\hypertarget{41-des-validation}{%
\subsection{4.1 DES validation}\label{41-des-validation}}

The reconstructed DES was first validated against the deterministic
baseline (Cf0, single shift S=1) reported in Garrido-Rios (2017, Table
6.10). Using the official thesis basis of 8,064 hours per year (48
operational weeks), the simulation delivers 733,621 rations per year on
average, compared with the thesis reference of 767,592 rations, yielding
a relative gap of $-4.43\%$. This is comfortably within the accepted
$\pm 15\%$ validation threshold and supports the structural fidelity of
the rebuilt DES. A secondary Gregorian-basis check (8,760 hours per
year) produces 796,940 rations per year ($+3.82\%$ relative gap),
confirming that the discrepancy is driven by year-basis conventions
rather than model error.

Under deterministic conditions with risks disabled, the simulation
produces 734,458 rations per year and delivers 733,621, with a fill rate
of 99.3\% and only 41 backorders accumulated over the 20-year horizon.
When stochastic risks are enabled at the current level, average annual
delivery falls to 677,750 rations ($-7.6\%$ relative to the
deterministic baseline), the fill rate drops to 68.3\%, and total
backorders rise to 1,825. Under the increased risk configuration used
for the benchmark, average annual delivery decreases further to 549,250
rations, fill rate falls to 45.6\%, and total backorders increase to
3,132. The DES therefore shows a clear transition from a well-calibrated
nominal system to a fragile network once disruption intensity escalates.

\hypertarget{42-track-a}{%
\subsection{4.2 Track A: thesis-faithful benchmark (negative
result)}\label{42-track-a}}

Under the thesis-faithful Track A contract, no RL configuration
outperforms the strongest static baseline. PPO trained with
ReT\_seq\_v1 ($\kappa = 0.20$) over 500,000 timesteps across 5 seeds
reaches a cumulative fill rate of 0.788, falling short of static S2 at
0.792 by 0.4 percentage points. PPO trained with the linear control\_v1
reward achieves fill rate 0.782, losing by 1.0 percentage point.
RecurrentPPO (LSTM-based) trained with control\_v1 performs worse still
at fill rate 0.751, losing by 4.1 percentage points. A tuned heuristic
baseline reaches 0.795, approximately matching S2 but not materially
exceeding it.

The negative result is not an algorithmic failure in the conventional
sense. A design-of-experiments analysis reveals that the Track A action
space exhibits strongly asymmetric sensitivity: destructive inventory
settings reduce reward by 44--65\% and can collapse fill rate, while the
positive headroom above S2-neutral inventory control is limited to
approximately 1\% in reward. The structural explanation is the
downstream distribution bottleneck: Op10--Op12 transport throughput
($\approx$2,500 rations/day) approximately equals demand, so extra
upstream production capacity from S3 yields sharply diminishing returns.
Because Track A does not expose downstream dispatch control to the
agent, the active constraint lies outside the controllable interface.

\hypertarget{43-track-b}{%
\subsection{4.3 Track B: extended action space (positive
result)}\label{43-track-b}}

Track B extends the action space from 5 to 7 dimensions by adding
downstream dispatch multipliers at Op10 and Op12, while preserving the
DES backbone and reward family. Table~\ref{tab:track_b_main} presents
the Track B results from the production benchmark (500,000 timesteps, 5
seeds $\times$ 20 evaluation episodes).

\begin{table}[htbp]
\centering
\caption{Track B main results (increased risk, 500k $\times$ 5 seeds). CI95 in brackets.}
\label{tab:track_b_main}
\begin{tabular}{lcccc}
\toprule
Policy & Fill Rate & Order-level ReT & Shift Mix (S1/S2/S3) & Autotomy\% \\
\midrule
\textbf{PPO} & \textbf{1.000} [1.000, 1.000] & \textbf{0.950} [0.946, 0.954] & 77/17/6 & 96.8 \\
RecurrentPPO & 1.000 [1.000, 1.000] & 0.949 [0.945, 0.953] & 78/14/7 & 96.6 \\
S1 ($d{=}1.0$) & 0.830 [0.822, 0.837] & 0.332 & 100/0/0 & 4.0 \\
S2 ($d{=}1.0$) & 0.958 [0.952, 0.964] & 0.469 & 0/100/0 & 14.7 \\
S2 ($d{=}1.5$) & 0.990 [0.983, 0.996] & 0.479 & 0/100/0 & 15.4 \\
S2 ($d{=}2.0$) & 0.985 [0.977, 0.993] & 0.463 & 0/100/0 & 12.7 \\
S3 ($d{=}1.0$) & 0.960 [0.952, 0.969] & 0.505 & 0/0/100 & 15.1 \\
S3 ($d{=}2.0$) & 0.985 [0.979, 0.991] & 0.449 & 0/0/100 & 10.6 \\
\bottomrule
\end{tabular}
\end{table}

PPO achieves fill rate 1.000 (zero backorders) and order-level
resilience 0.950. The best static policy (S2 with $d{=}1.5$) reaches
fill rate 0.990; PPO exceeds it by 1.0 percentage points. RecurrentPPO
matches PPO's performance, confirming that recurrence does not add value
when the action space is already aligned with the bottleneck.

The learned policy's shift utilization is 77\% S1 / 17\% S2 / 6\%
S3---a marked contrast to static S2 (100\% S2) and static S3 (100\%
S3). PPO learned to operate at the lowest assembly cost (S1) for most of
the episode while activating higher shifts selectively. The mean
downstream dispatch multipliers are 1.36 (Op10) and 1.50 (Op12), with
high dispatch ($\geq 1.9$) occurring on only 17\% of steps. This
represents a cost-efficient strategy that no static policy achieves:
PPO uses approximately 18,500 assembly hours versus $\approx$43,700 for
S3($d{=}2.0$)---a 57\% reduction---while delivering higher fill rate.

\hypertarget{44-causal-ablation}{%
\subsection{4.4 Causal ablation: action-space decomposition}\label{44-causal-ablation}}

To isolate the effect of downstream control, we conduct a production-grade
ablation (500,000 timesteps, 5 seeds) holding all factors constant except the
action contract. All conditions use observation version v7 (46 dimensions),
reward ReT\_seq\_v1 ($\kappa = 0.20$), and risk level adaptive\_benchmark\_v2.
Three configurations are compared: \emph{joint} (full 7D), \emph{shift\_only}
(downstream frozen at $1.25\times$), and \emph{downstream\_only} (shift frozen
at S2).

\begin{table}[htbp]
\centering
\caption{Causal ablation (increased risk, 500k $\times$ 5 seeds).}
\label{tab:ablation}
\begin{tabular}{lcccc}
\toprule
Action Config & Fill Rate & Order-level ReT & Assembly Hours & Beats Best Static? \\
\midrule
\textbf{Joint (7D)} & \textbf{1.000} & 0.948 & 18,529 & Yes \\
Downstream-only & 1.000 & 0.953 & 33,322 & Yes \\
Shift-only & 0.953 & 0.686 & 26,295 & No \\
Best static (S2, $d{=}1.5$) & 0.990 & 0.479 & $\approx$29,000 & --- \\
\bottomrule
\end{tabular}
\end{table}

The results establish a clear causal story. Without downstream control
(\emph{shift\_only}), PPO cannot beat the best static baseline---reproducing
the Track A failure even with the enriched v7 observation space. With
downstream control alone (\emph{downstream\_only}), PPO achieves fill rate
1.000, confirming that downstream dispatch is the necessary and sufficient
action dimension for fill improvement. The \emph{joint} configuration adds
cost efficiency: 57\% fewer assembly hours than the best static, achieved by
combining S1-dominant shifts with selective downstream dispatch. This
decomposition provides the strongest causal evidence in the paper: the
downstream control dimensions, not algorithmic sophistication or reward
engineering, determine the benchmark outcome.

\hypertarget{45-reward-sweep}{%
\subsection{4.5 Reward insensitivity under Track B}\label{45-reward-sweep}}

To assess the robustness of the Track B result to reward design, we train
PPO under 7 distinct reward formulations (500,000 timesteps, 5 seeds each),
all using the same Track B action contract and observation space.

\begin{table}[htbp]
\centering
\caption{Reward sweep (Track B, PPO, 500k $\times$ 5 seeds).}
\label{tab:reward_sweep}
\begin{tabular}{lccc}
\toprule
Reward Mode & Fill Rate & Order-level ReT & Converges? \\
\midrule
ReT\_cd\_v1 & 1.000 & 0.954 & Yes \\
control\_v1 & 1.000 & 0.953 & Yes \\
ReT\_seq\_v1 & 1.000 & 0.951 & Yes \\
ReT\_garrido2024\_train & 1.000 & 0.948 & Yes \\
ReT\_unified\_v1 & 1.000 & 0.945 & Yes \\
\midrule
ReT\_corrected & 0.843 & 0.269 & No (S1 collapse) \\
ReT\_thesis & 0.836 & 0.258 & No (S1 collapse) \\
\bottomrule
\end{tabular}
\end{table}

Five of seven reward formulations converge to the same strong policy region
(fill $\geq 0.999$, ReT $\geq 0.945$). The two failures
(ReT\_corrected and ReT\_thesis) share a common structure: both implement
piecewise step-level approximations of the thesis resilience function that
produce sharp gradients encouraging S1 collapse. The convergence of five
diverse rewards---including linear (control\_v1), sequential C-D
(ReT\_seq\_v1), and intermediate-variable C-D
(ReT\_garrido2024\_train)---demonstrates that when the action space covers
the active bottleneck, the specific choice of training reward becomes
almost irrelevant among well-calibrated formulations. This contrasts sharply
with Track A, where reward choice appeared critical but could not overcome
structural limitations.

\hypertarget{46-cross-scenario}{%
\subsection{4.6 Cross-scenario robustness}\label{46-cross-scenario}}

We evaluate the frozen PPO policy (trained under adaptive\_benchmark\_v2)
across four risk severity levels to characterize the boundary of RL
effectiveness.

\begin{table}[htbp]
\centering
\caption{Cross-scenario evaluation (frozen PPO, 5 seeds $\times$ 20 episodes).}
\label{tab:cross_scenario}
\begin{tabular}{lcccccc}
\toprule
Risk Level & PPO Fill & PPO ReT & Best Static Fill & Best Static ReT & $\Delta$ Fill (pp) \\
\midrule
Current & 1.000 & 0.915 & 0.9997 & 0.733 & +0.0 \\
Increased & 0.993 & 0.748 & 0.901 & 0.202 & +9.2 \\
Severe & 0.966 & 0.424 & 0.656 & 0.162 & +31.0 \\
Severe extended & 0.558 & 0.147 & 0.282 & 0.152 & +27.6 \\
\bottomrule
\end{tabular}
\end{table}

PPO's advantage grows monotonically with disruption severity from current
(+0.0pp) through severe (+31.0pp). Under severe\_extended conditions,
PPO still wins in fill rate (+27.6pp) but loses marginally in order-level
ReT (0.147 vs.\ 0.152 for the best static). This represents the
operational boundary: under extreme stress, the policy enters permanent
recovery mode (99.8\% of orders classified as recovery), and autotomy drops
from 96.8\% (increased) to 0.1\%. The degradation curve provides honest
evidence that PPO is not invincible---it has a definable failure regime.

\hypertarget{47-forecast-sensitivity}{%
\subsection{4.7 Forecast sensitivity}\label{47-forecast-sensitivity}}

To determine whether PPO exploits the numerical content of risk forecasts
in the observation vector, we evaluate the frozen policy under three
conditions: full (unmodified), scrambled (randomly permuted across risk
types), and zeroed (all forecasts set to 0).

\begin{table}[htbp]
\centering
\caption{Forecast sensitivity (frozen PPO, 5 seeds $\times$ 20 episodes).}
\label{tab:forecast}
\begin{tabular}{lccc}
\toprule
Condition & Fill Rate & Order-level ReT & Autotomy\% \\
\midrule
Full forecasts & 1.000 & 0.950 & 96.8 \\
Scrambled & 1.000 & 0.951 & 96.8 \\
Zeroed & 1.000 & 0.913 & 90.6 \\
\bottomrule
\end{tabular}
\end{table}

Scrambling forecasts produces no degradation (ReT 0.951 vs.\ 0.950),
confirming that PPO does not use the numerical content of risk
predictions. Zeroing forecasts causes mild degradation in ReT
($-3.7$pp) and autotomy ($-6.2$pp) but no fill degradation, suggesting
this reflects out-of-distribution observation effects rather than
forecast dependence. A regime-conditioned analysis across 5 seeds shows
that PPO escalates shift usage 2.6$\times$ during pre-disruption
regimes compared to nominal periods ($P(S>1|\text{pre\_disruption}) =
0.323$ vs.\ $P(S>1|\text{nominal}) = 0.126$), confirming
regime-responsive behavior. The policy is therefore better characterized
as a \emph{regime-responsive adaptive controller} than as an
anticipatory predictor.

\hypertarget{48-summary}{%
\subsection{4.8 Summary}\label{48-summary}}

The experimental evidence forms a coherent narrative across seven
result dimensions: Track A demonstrates a structural limitation (Section
4.2); Track B demonstrates that a minimal action-space repair overcomes
it (Section 4.3); the causal ablation localizes the effect to downstream
control (Section 4.4); the reward sweep shows insensitivity to reward
choice when the action space is aligned (Section 4.5); cross-scenario
evaluation establishes the boundary of effectiveness (Section 4.6); and
forecast sensitivity reveals a regime-responsive policy mechanism
(Section 4.7). The central finding is not that RL is universally
effective, but that its effectiveness depends critically on whether the
controllable interface reaches the binding operational constraint.
