\section{Results}\label{sec:results}

\subsection{Comparator scope reverses the apparent adaptive headline}

Against a 147-cell frontier that varied shift and downstream dispatch while holding upstream
controls fixed, ten PPO checkpoints produced a crossed held-out ReT difference of $+0.000486$.
This comparison was stable but did not grant the static policy the learner's complete action
contract. The frozen same-contract challenge selected a constant using calibration tapes and
evaluated it on 60 fresh tapes. Full-contract static ReT was $0.005906366$ versus $0.005888317$
for PPO; PPO minus static was $-0.000018049$, with two-way CI95
$[-0.000028615,-0.000008087]$. A small within-learner dispatch increment remained positive,
but it did not restore superiority over static control. The restricted comparison therefore
diagnoses comparator-family sensitivity rather than adaptive advantage.

\subsection{Decision families fail at different rungs of the ladder}

The boundary programs reveal distinct failure mechanisms (Table~\ref{tab:eligibility}).
DRA-1 created physically live allocation states, but balanced-prefix exact branching found
negligible and sparse dynamic oracle value. DRA-2b was qualitatively different: the restricted
14-day open-loop oracle beat a budget-feasible static comparator by $0.02212$ ReT, CI95
$[0.01725,0.02734]$, and both HOLD and DISPATCH were supported in substantial state subsets.
It nevertheless failed its preregistered service-practicality and horizon-sufficiency gates.
This was evidence of clairvoyant headroom, not deployable policy value.

Program E directly tested conversion using new data and deployable observations. Ten
MaskablePPO seeds were trained for 200,000 steps each, alongside an interpretable tree and
heuristics. Validation conversion efficiency was $-0.00438$ and no PPO seed produced a
positive effect over the same-contract static bar. The tree attained high first-action
classification accuracy in development but lost in sequential rollout. Thus a learner was
trained under a legitimate headroom-motivated contract; its null cannot be attributed to a
failure to attempt RL.

Program F imposed exactly two readiness tokens across maintenance, transport protection, and
reserve actions, thereby eliminating resource-purchase explanations by construction. Across
the preregistered phase diagram, observable-tree rollout conversion passed in $0/24$ cells.
Several cells displayed oracle signal, but every admissible tree rollout delta was negative.
The program therefore stopped before learner training.

\begin{table}[htbp]
\centering
\caption{Eligibility ladder across the main terminal studies. A dash means the program stopped
before that rung could support a claim.}
\label{tab:eligibility}
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lccccc}
\toprule
Study & Physical & Ranking & PI headroom & Observable rollout & Learner win \\
\midrule
Track A / Program L & yes & weak/limited & small & no & no \\
DRA-1 allocation & yes & insufficient & negligible & -- & -- \\
DRA-2b convoy & yes & yes & $+0.02212$ & not authorized & -- \\
Program E convoy & yes & yes & material prior & no & no (0/10) \\
Program F portfolio & yes & mixed & cell-dependent & 0/24 gates & -- \\
Program G spatial & yes & yes & diagnostic & no & -- \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Program G: corrected terminal spatial audit}

Program G initially appeared promising under a ration-mass proxy, motivating a stricter
order-ledger audit. Two defects were corrected before the terminal test: the six-day operating
week had been compressed to 144 hours instead of retaining a 168-hour calendar, and two
runners scored the current order before updating cumulative backlog and unattended counters.
The corrective protocol was committed before opening 200 new calibration tapes and 400 locked
test tapes. It also added quantity-weighted ReT and marked horizon-pending orders lost.

The best periodic same-contract policy selected on calibration was blind alternation, ABAB.
No observable policy passed any substantive joint gate. Cover minus ABAB yielded order ReT
$-0.02317$ (CI95 $[-0.02816,-0.01893]$), quantity-weighted ReT $-0.00695$
(CI95 $[-0.00869,-0.00548]$), $3.51$ fewer attended orders, worst-CSSU fill $-0.1734$,
and 3,370 additional unfulfilled rations at the horizon. MPC also failed every guardrail; the
service and ReT depth-3 trees collapsed to the same tested behavior as cover. Equal action
rights rule out a resource-advantage explanation.

\begin{table}[htbp]
\centering
\caption{Program G terminal locked test: observable cover policy minus frozen ABAB
($n=400$ paired tapes). Positive is desirable except for unfulfilled rations.}
\label{tab:programg}
\small
\begin{tabular}{lrr}
\toprule
Endpoint & Mean delta & CI95 \\
\midrule
Canonical ReT per order & $-0.02317$ & $[-0.02816,-0.01893]$ \\
Quantity-weighted ReT & $-0.00695$ & $[-0.00869,-0.00548]$ \\
Attended orders & $-3.51$ & $[-4.15,-2.95]$ \\
Worst-CSSU fill & $-0.1734$ & $[-0.2033,-0.1466]$ \\
Unfulfilled rations & $+3370$ & $[+2831,+3971]$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Resilience functionals induce different rankings}

The corrected triangulation refutes a simple order-versus-mass explanation: both order ReT and
quantity-weighted ReT favor ABAB, as do attended orders and worst-CSSU fill. A
Cobb--Douglas-inspired aggregate instead favors cover slightly. Its capacity and cost terms are
constant in this contract, and its remaining inventory--backlog--time aggregation does not
preserve spatial fairness. The disagreement therefore documents a normative preference for
concentration, not hidden operational superiority. We retain the metric reversal as a
methodological result: adaptive-control eligibility is not invariant to the resilience
functional, and a multiplicative aggregate can reward a policy that worsens continuity and the
worst-served node.

\subsection{Program H: belief-state audit}

The final spatial study asked whether Program G failed because its policy classes were weak or
because its information was insufficient. An O0 Bayes filter improved tempo log loss from
1.0016 to 0.8568, confirming information in the signal history. Nevertheless, regret fitted-Q
improved order ReT by only 0.00225 (CI95 $[-0.00021,0.00460]$), was favorable on 19\% of
tapes, and converted 13.7\% of the perfect-information gap. It also attended fewer orders and
worsened worst-CSSU fill. Belief-MPC and point rollout were negative. No policy passed the
learner gate, so the virgin RL universe remained unopened. The full-tape relaxation ceiling
remained +0.01641 (CI95 $[0.01397,0.01886]$); this loose ceiling prevents a formal
impossibility claim. The conclusion is a tested-policy null with unresolved optimal observable
value.
